### 2011-06-17 Documentation/Article.tex
Working on Intro

%
% Complete documentation on the extended LaTeX markup used for Insight
% documentation is available in ``Documenting Insight'', which is part
% of the standard documentation for Insight.  It may be found online
% at:
%
%     http://www.itk.org/

\documentclass{InsightArticle}

\usepackage[dvips]{graphicx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  hyperref should be the last package to be loaded.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[dvips,
bookmarks,
bookmarksopen,
backref,
colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue},
]{hyperref}

\usepackage{url}

%  This is a template for Papers to the Insight Journal. 
%  It is comparable to a technical report format.

% The title should be descriptive enough for people to be able to find
% the relevant document.
\title{CTest Integration of Sikuli Automated GUI Testing}

% 
% NOTE: This is the last number of the "handle" URL that 
% The Insight Journal assigns to your paper as part of the
% submission process. Please replace the number "1338" with
% the actual handle number that you get assigned.
%
\newcommand{\IJhandlerIDnumber}{3196}

% Increment the release number whenever significant changes are made.
% The author and/or editor can define 'significant' however they like.
\release{0.00}

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{Evan Schwab, Lydie Souhait, Nicolas Rannou, Kishore Mosaliganti\\
Arnaud Gelas, Sean Megason}
\authoraddress{Harvard Medical School, Megason lab}

\begin{document}

% Add hyperlink to the web location and license of the paper.
\IJhandlefooter{\IJhandlerIDnumber}

\ifpdf
\else
   %
   % Commands for including Graphics when using latex
   %
   \DeclareGraphicsExtensions{.eps,.jpg,.gif,.tiff,.bmp,.png}
   \DeclareGraphicsRule{.jpg}{eps}{.jpg.bb}{`convert #1 eps:-}
   \DeclareGraphicsRule{.gif}{eps}{.gif.bb}{`convert #1 eps:-}
   \DeclareGraphicsRule{.tiff}{eps}{.tiff.bb}{`convert #1 eps:-}
   \DeclareGraphicsRule{.bmp}{eps}{.bmp.bb}{`convert #1 eps:-}
   \DeclareGraphicsRule{.png}{eps}{.png.bb}{`convert #1 eps:-}
\fi

\maketitle

\ifhtml
\chapter*{Front Matter\label{front}}
\fi

\begin{abstract}
\noindent

\end{abstract}

\IJhandlenote{\IJhandlerIDnumber}

\tableofcontents
% ------------------------------------------------------------------------
\section{Introduction}
GUI testing is a key aspect of software development. Having a well defined testing protocol is very
important in maintaining a productive flow, especially on the eve of a new release.  The number of units
to test is greatly compounded as the software becomes more complex and keeping up with the complexity can be
immensely time consuming.  We collected a list of over 400 different but very repetitive GUI tests which
involved endless clicking and checking.  This task took around 5 hours to complete each time a new release was ready.
Then debugging would push release dates back days or even weeks.   
% ------------------------------------------------------------------------
\section{Sikuli}

% Feel free to change section titles
\subsection{What is sikuli?}

Sikuli~\cite{Sikuli:Documentation,Sikuli:Website,Yeh:2009:Sikuli} is a visual
technology to automate and test graphical user interfaces (GUI) using images
(screenshots). Sikuli includes \emph{Sikuli Script} and \emph{Sikuli IDE}.\\

\emph{Sikuli Script} automates anything you see on the screen without internal
API's support. Sikuli Script is a programming language that uses screen shot
images as variables and objects in order to automate graphical user interface
functions. The Sikuli Script is built on a Jython (Python for Java platform)
library which uses Python syntax in addition to a number of special Sikuli
functions for acquiring and handling screen shot images and performing mouse,
keyboard actions.\\

In the \emph{Sikuli IDE}, the user can write scripts that include screen shot
thumbnails so they can visually track the functions of their code. The most
useful functions from Sikuli scripts are conveniently found as
buttons on the IDE. When using the IDE, Sikuli then saves a script in a
directory with a unique extension .sikuli. The directory consists of a python
script file (script file to be used by sikuli), an html file, and a list of all
the screenshot images used in the script.

\subsection{Simple Sikuli example} % not related to gofigure...

Figure~\ref{fig:SimpleExample} shows the Sikuli-IDE for version 0.10.2. On the
left-hand side menu is a list of commonly used Sikuli commands
like \emph{find}(), \emph{click}(), and \emph{type}(). The picture of
the camera within the parentheses indicates the screenshot input variable.
Clicking on any of these menu functions will automatically set up the Sikuli
screen shot mode, whereby the Sikuli-IDE disappears and what is on the screen at
that moment is available to capture by selecting a rectangular region of
interest. The function will then appear in the scripting domain with the snap
shot in parentheses as shown in Figure~\ref{fig:SimpleExample}. The user may
also type in the function and invoke the screen shot mode by clicking on the
camera in the top menu or by using a keyboard shortcut. The simple Sikuli script
example in Figure~\ref{fig:SimpleExample} will open gedit Text Editor, type a
message, and save the file.

\begin{figure}[htbp]
 \centering
 \includegraphics[width=0.99\textwidth]{Images/SimpleSikuliExample.png}
 % SimpleSikuliExample.png: 1280x1024 pixel, 72dpi, 45.16x36.12 cm, bb=0 0 1280 1024
 \caption{Simple Sikuli Script Example in Sikuli-IDE}
 \label{fig:SimpleExample}
\end{figure}

% ------------------------------------------------------------------------
\section{Example}

The Sikuli script in Figure~\ref{fig:Gofigure2Example} shows a fragment of a
unit test for the GoFigure2 software~\cite{GoFigure2:Website}. The unit being
tested here is the Toolbar menu which consists of different modes used to
analyze data set plots. In total there are four modes which are represented in
the first line as the vector Toolbar and for each of these modes we want to test
their expected mouse functions. For instance, if we are in the zoom Toolbar
mode then we want the data set plot to zoom in when the user scrolls the mouse
wheel forward. Likewise, if we are in pan mode then we want the plot to move to
the left if the user holds down the left mouse button and moves the mouse to the
left.  Also importantly, we want to see that these two mouse actions are unique
for their respective Toolbar modes; the test will include an assertion that
scrolling forward will zoom in when using zoom mode and will do nothing when
using pan mode.\\

This breakdown of modes and actions lends itself to a nested for-loop scripting
approach. By representing these variables as a vector the software developers
can easily add new modes or actions to each list to be tested, when the software
is updated. Similarly, the tester can easily replace any of the screenshot
images with images of updated GUI icons.\\

In addition to representing the modes and actions as the vectors Toolbar and
Mouse, respectively, we have included the different data set plots in the
vector ViewRegion. These images represent the \emph{xy-}, \emph{xz-},
\emph{yz-}, and \emph{xyz-}views of zebrafish microscopy data taken in the
Megason Lab. After capturing these images we can then use them to locate their
initial pixel coordinates within the GoFigure2 GUI using the \emph{find}()
function. These points become our reference positions to calculate directional
movement.  For example, when testing the pan mode, we mark the initial position
of each image, use the \emph{mouseMove}() function to a chosen location, and
then use \emph{find}() to obtain the new coordinates. We can then assert if the
image panned as it should have. Similarly for zoom, we can first find the area
of the image and then find the area of the zoomed in image.  To test the result
when the user clicks in a certain mode, we've made use of Sikuli's \emph{assert
exists}() function to ascertain the existence of our desired results (see
Figure~\ref{fig:Gofigure2Example}).\\

This one GUI test is then combined with other GoFigure2 tests in succession and
the Sikuli Unit Testing CTest integration will identify which tests pass and
which fail and when. Many of the units need to be tested in a certain order.
For example, in order to test the Toolbar modes on the data sets, one first
needs to test the software's data import functionality.

\begin{figure}[htp]
 \centering
 \includegraphics[width=0.99\textwidth]{Images/Gofigure2Example.png}
 % Gofigure2Example.png: 667x826 pixel, 72dpi, 23.53x29.14 cm, bb=0 0 667 826
 \caption{Sikuli Script for Gofigure2 GUI Testing}
 \label{fig:Gofigure2Example}
\end{figure}

% ------------------------------------------------------------------------
\section{Integration with CTest}

% I'll write it this part
One of Sikuli's primary uses is the automation of GUI testing for software
developers. For GUI testing, it is useful to compartmentalize different software
functions and test each unit individually.\\

The newest versions of Sikuli come equipped with a Unit Testing feature based
on JUnit~\cite{JUnit:Website}. Unit Testing can be called using the IDE or by
command line and enables the user to write multiple functions in a script which
are run individually. At the end of the run, the Sikuli output will tell the
user how many tests ran successfully and which ones failed and where.\\

Here we connect this Unit Testing feature to CTest, this allows us to report
results on public dashboard using CDash. To this end, we developed cmake
functions to discover where Sikuli is installed and special function to add
sikuli test in the list of test for CTest.

Here is a typical CMakeLists.txt file, to be able to use such functionalities:

\begin{verbatim}
# first let's include CTest and enable testing

include( CTest )
enable_testing()

# Then let's
#   - figure out where Sikuli is installed
#   - define function to add Sikuli tests

find_package( Sikuli )

# here we add a sikuli test
add_sikuli_test( ThunderbirdSikuliTest test.sikuli )
\end{verbatim}

Users can either maintain an image library, either keep images in the
corresponding sikuli script directory (default). If one use an image library,
the corresponding path must be set during CMake time in
the SIKULI\_IMAGE\_LIBRARY\_DIR variable.

% ------------------------------------------------------------------------
\section{Conclusion}

\clearpage

\bibliographystyle{plain}
\bibliography{InsightJournal,biblio}

\end{document}